---
title: "Fire model simulations"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

## setup

```{r setup, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=TRUE}
needed_packages <- c(
    "tidyverse"
  , "lubridate"
  , "nlrx"
  , "tictoc"
  , "future"
  , "poweRlaw"
  , "future.apply"
)

lapply(needed_packages, function(x) { if(!require(x,character.only = TRUE)) install.packages(x)} )

theme_set(theme_bw())
source("R/functions.r")

# Data Path
#
if( Sys.info()['nodename'] =="ls-pro") {
  
  data_path <- "~/Academicos/GitProjects/AustraliaTippingPoint/Data"
} else if(Sys.info()['nodename'] =="biologia2018") {                     # Server UNGS
  data_path <- "~/GitProjects/AustraliaTippingPoint/Data"

}

#
# NetLogo 
# Setup for nlrx
#

if( Sys.info()['nodename'] =="ls-pro") {
  simfolder <- "/home/leonardo/Academicos/GitProjects/fireNL"

} else if(Sys.info()['nodename'] =="biologia2018") {                     # Server UNGS
  simfolder <- "/home/leonardo/GitProjects/fireNL"
}


# Unix default NetLogo installation path (adjust to your needs!):
#
netlogopath <- file.path("/home/leonardo/NetLogo")
modelpath <- file.path(simfolder, "DynamicFireForest.nlogo")
outpath <- file.path(simfolder,"Simulations")

# If not defined set the JAVA version of your local 
if(Sys.getenv("JAVA_HOME")==""){
  Sys.setenv(JAVA_HOME = "/usr/lib/jvm/java-11-openjdk-amd64")
  ## "/usr/lib/jvm/java-8-oracle"
}

nl <- nl(nlversion = "6.2",
         nlpath = netlogopath,
         modelpath = modelpath,
         jvmmem = 2048)



```



## Simulations with patch size output for wordl_width = 450 for different theta


```{r sim_ff449cluster40, echo=FALSE, tidy=TRUE, message=FALSE, warning=FALSE, eval=FALSE}


disp_best <- c(1.01, 65.72)
fire_best <- c(2e-7,2e-6,2e-5)

eval_times <- as.numeric(lapply(0:39, function(x){ymd("2021-12-31") + years(x) - ymd("2001-01-01") }))
ymd("2001-01-01") + days(eval_times)

#nl@experiment <- experiment(expname="Amazon20years449sp",

nl@experiment <- experiment(expname="noseason40years449theta25_2500",
                            outpath=outpath,
                            repetition=1,
                            tickmetrics="true",
                            idsetup="setup",
                            idgo="go",
                            #idrunnum="nlrx-experiment",
                            runtime=0,
                            evalticks=eval_times,          # ymd("2011-12-31")  - ymd("1980-11-01") + 0:19
                            metrics=c("Date", "burned-by-month","active-burned","Percent-forest","fire-probability","burned-clusters 365","median-fire-interval" ),
                            variables = list("forest-dispersal-distance" = list(values=disp_best),
                                             "Fire-probability" = list(values=fire_best)),
                            constants = list("world-width" = 449,
                                             "world-height" = 449,
                                               "fire-prob-filename"="\"Data/Estimated_bF.ppp\"",  # No data from file
                                               "Save-view" =  "false",
                                               "video" = "false",
                                               "end-simulation"= 21915,              # 40 years 
                                               "use-fire-prob-se" = "false",
                                               "Initial-forest-density" = 0.3,
                                               "Periodicity" = "false",
                                               "Forest-growth" = 2000,
                                               "eval-burned-clusters"=paste0("\"[",paste(eval_times,collapse=" "),"]\"")
                                             ))


#
# Run 10 times this do not set the run variable for output 
#
# nl@simdesign <- simdesign_simple(nl=nl,
#                                nseeds=1)

nl@simdesign <- simdesign_ff(nl=nl,
                               nseeds=10)

# run in Paralell 
#
plan(multisession)
tic()
#results <- run_nl_all(nl)
results <- run_nl_all(nl,split = 6)
toc()
plan(sequential)
names(results)

#
# Write the output 
#
setsim(nl, "simoutput") <- results 
write_simoutput(nl)

```

## Calculate the fires cluster size distribution 

```{r comp_ff449clusterSize, echo=FALSE, tidy=TRUE, message=FALSE, warning=FALSE, eval=TRUE}

# Read previously saved fitted exponents file
#
if(file.exists(file.path("Simulations", "fittedExponentsFireNL40years.rds"))) {
  pexp <- readRDS(file.path("Simulations", "fittedExponentsFireNL40years.rds")) 
  pexp %>% distinct(world_width,sim_type)
  
} else {
  pexp <- tibble()
  
}

sim <- read_netlogo_simul(file.path(outpath,"noseason40years449theta25_2500_ff.csv"),skip=0)

plan(multisession)
p_df <- future_lapply( 1:nrow(sim) , function(i){
  df <- netlogo_evaluate_patch_distr(sim$burned_clusters_365[i]) %>%
    mutate(world_width=sim$world_width[i],Fire_probability = sim$Fire_probability[i],forest_dispersal_distance=sim$forest_dispersal_distance[i],
           Forest_growth=sim$Forest_growth[i],Date=sim$Date[i],random_seed=sim$random_seed[i], Percent_forest=sim$Percent_forest[i], 
           median_fire_interval= sim$median_fire_interval[i]) %>%
    dplyr::select(world_width,Fire_probability,forest_dispersal_distance,Forest_growth,Date,random_seed, everything()) %>% mutate(sim_type="fixed") %>% dplyr::select(-date)
})
pexp <- bind_rows(pexp,p_df)
pexp <- pexp %>% fill(range)%>% drop_na(type)
plan(sequential)

saveRDS(pexp, file.path("Simulations", "fittedExponentsFireNL40years.rds"))

```


## Simulations with patch size output for wordl_width = 450 for different theta and seasonality


```{r sim_ff449clusterSeason_ini03, echo=FALSE, tidy=TRUE, message=FALSE, warning=FALSE, eval=FALSE}


disp_best <- c(1.01, 65.72)
fire_best <- c(2e-7,2e-6,2e-5)

eval_times <- as.numeric(lapply(0:39, function(x){ymd("2021-12-31") + years(x) - ymd("2001-01-01") }))
ymd("2001-01-01") + days(eval_times)

nl@experiment <- experiment(expname="season40years449theta25_2500",
                            outpath=outpath,
                            repetition=1,
                            tickmetrics="true",
                            idsetup="setup",
                            idgo="go",
                            #idrunnum="nlrx-experiment",
                            runtime=0,
                            evalticks=eval_times,          # ymd("2011-12-31")  - ymd("1980-11-01") + 0:19
                            metrics=c("Date", "burned-by-month","active-burned","Percent-forest","fire-probability","burned-clusters 365","median-fire-interval" ),
                            variables = list("forest-dispersal-distance" = list(values=disp_best),
                                             "Fire-probability" = list(values=fire_best)),
                            constants = list("world-width" = 449,
                                             "world-height" = 449,
                                               "fire-prob-filename"="\"Data/Estimated_bF.ppp\"",  # No data from file
                                               "Save-view" =  "false",
                                               "video" = "false",
                                               "end-simulation"= 21915,              # 40 years 
                                               "use-fire-prob-se" = "false",
                                               "Initial-forest-density" = 0.3,
                                               "Periodicity" = "true",
                                               "increase-fire-prob-seasonality" = 10,
                                               "days-fire-season" = 90,
                                               "Forest-growth" = 2000,
                                               "eval-burned-clusters"=paste0("\"[",paste(eval_times,collapse=" "),"]\"")
                                             ))


#
# Run 10 times this do not set the run variable for output 
#
# nl@simdesign <- simdesign_simple(nl=nl,
#                                nseeds=1)

nl@simdesign <- simdesign_ff(nl=nl,
                               nseeds=10)

# run in Paralell 
#
plan(multisession)
tic()
#results <- run_nl_all(nl)
results <- run_nl_all(nl,split = 6)
toc()
plan(sequential)
names(results)

#
# Write the output 
#
setsim(nl, "simoutput") <- results 
write_simoutput(nl)

```



## Cluster distribution comparison 

* Calculate the fires cluster size distribution 

```{r dist_season_ff449clusterSize, echo=FALSE, tidy=TRUE, message=FALSE, warning=FALSE, eval=TRUE}

#
# Read previously saved fitted exponents file
#
if(file.exists(file.path("Simulations", "fittedExponentsFireNL40years.rds"))) {
  pexp <- readRDS(file.path("Simulations", "fittedExponentsFireNL40years.rds")) 
  pexp %>% distinct(world_width,sim_type,forest_dispersal_distance,Fire_probability)
  
} else {
  pexp <- tibble()
  
}

sim <- read_netlogo_simul(file.path(outpath,"season40years449theta25_2500_ff.csv"),skip=0)

plan(multisession)
p_df <- future_lapply( 1:nrow(sim) , function(i){
  df <- netlogo_evaluate_patch_distr(sim$burned_clusters_365[i]) %>%
    mutate(world_width=sim$world_width[i],Fire_probability = sim$Fire_probability[i],forest_dispersal_distance=sim$forest_dispersal_distance[i],
           Forest_growth=sim$Forest_growth[i],Date=sim$Date[i],random_seed=sim$random_seed[i], Percent_forest=sim$Percent_forest[i], 
           median_fire_interval= sim$median_fire_interval[i]) %>%
    dplyr::select(world_width,Fire_probability,forest_dispersal_distance,Forest_growth,Date,random_seed, everything()) %>% mutate(sim_type="seasonal") %>% dplyr::select(-date)
})

pexp <- bind_rows(pexp,p_df)
pexp <- pexp %>% fill(range)%>% drop_na(type)
plan(sequential)

saveRDS(pexp, file.path("Simulations", "fittedExponentsFireNL40years.rds"))


```

## Plots 

* Calculate the fires cluster size distribution 

```{r plot_sim_ff449cluster, echo=FALSE, tidy=TRUE, message=FALSE, warning=FALSE, eval=FALSE}

#
# Read simulations based on fitted bF 
#

pexp <- readRDS(file.path("Simulations", "fittedExponentsFireNL40years.rds")) %>% mutate(tot_patch= tot_patch/(450*450)*100,max_patch=max_patch/(450*450)*100,num_patch=num_patch/(450*450)*100, year= year(Date), theta = 1/(Fire_probability * Forest_growth))


#
# Only the records with pl 
#
plot_pexp <- pexp %>% group_by(forest_dispersal_distance,theta,Date,random_seed) %>% mutate(deltaAIC = AICc - min(AICc)) %>%  filter( deltaAIC==0)

knitr::kable(plot_pexp %>% group_by(forest_dispersal_distance,theta,sim_type,type) %>% summarise(n = n(),range=mean(range),mean_patch=mean(mean_patch),expo=median(expo)) %>%  mutate(freq = n / sum(n),expo=if_else(type=="pl",expo,0)), digits = 2)



plot_pexp %>% group_by(forest_dispersal_distance,theta,sim_type,type) %>% summarise(n = n(),range=mean(range),mean_patch=mean(mean_patch),expo=median(expo)) %>%  mutate(freq = n / sum(n),expo=if_else(type=="pl",expo,0),theta=factor(theta),forest_dispersal_distance=round(forest_dispersal_distance)) %>% 
ggplot(aes(x=theta,y=freq,fill=type)) + 
  geom_bar(stat="identity", position=position_dodge()) + 
  scale_fill_viridis_d() + 
  xlab("") + ylab("") + theme_bw() + 
    theme(legend.position="bottom", legend.direction="horizontal") + 
    facet_wrap(vars(forest_dispersal_distance)) + geom_text(aes(label = round(range)), vjust = -0.2,position = position_dodge(.9))

# best_pexp <- pexp %>% group_by(forest_dispersal_distance,theta,Date,random_seed) %>% mutate(deltaAIC = AICc - min(AICc)) %>%  filter( deltaAIC==0 , nth(deltaAIC,2,order_by = deltaAIC) < 2) 
# 
# knitr::kable(best_pexp %>% group_by(forest_dispersal_distance,theta,sim_type,type) %>% summarise(n = n(),range=mean(range),mean_patch=mean(mean_patch)) %>%  mutate(freq = n / sum(n)), digits = 2)

pl_best_pexp <- plot_pexp %>% filter(type == "pl") %>% mutate( true_pl = if_else(exp_p_value<0.05, TRUE,FALSE)) # %>% mutate(relative_range=range/(350*350))

#
# Frequency of true 
#
knitr::kable(pl_best_pexp %>% group_by(forest_dispersal_distance,sim_type,theta,true_pl) %>% summarise(n = n(),iqr=IQR(expo),range=mean(range),expo=median(expo)) %>%  mutate(freq = n / sum(n)) %>% filter(true_pl==TRUE),digits=2)
pd <- position_dod(.2)
pl_best_pexp %>% group_by(sim_type,forest_dispersal_distance,theta) %>% summarize(iqr=IQR(expo),expo=median(expo)) %>% mutate(theta=factor(theta),forest_dispersal_distance=factor(round(forest_dispersal_distance))) %>% 
  ggplot( aes(x=theta, y=expo, colour=forest_dispersal_distance)) + 
    geom_errorbar(aes(ymin=expo-iqr, ymax=expo+iqr), width=.1, position=pd) +
    geom_point(position=pd)+ scale_color_viridis_d(name="Dispersal\nDistance") + 
    theme(legend.position = c(0.8, 0.8)) + ylab("Exponent") + xlab( expression(~theta)) + facet_wrap(~sim_type)
# + geom_hline(yintercept=2.29, linetype="dashed",  color = "red", size=1) 
ggsave("figure/FireNL_exponent_dispersal_theta_season.png",width=8,height=6,units="in",dpi=600)

pl_best_pexp %>% group_by(sim_type,forest_dispersal_distance,theta) %>% filter(true_pl==TRUE) %>% summarize(iqr=IQR(expo),expo=median(expo)) %>% mutate(theta=factor(theta),forest_dispersal_distance=factor(round(forest_dispersal_distance))) %>% 
  ggplot( aes(x=theta, y=expo, colour=forest_dispersal_distance)) + 
    geom_errorbar(aes(ymin=expo-iqr, ymax=expo+iqr), width=.1, position=pd) +
    geom_point(position=pd)+ scale_color_viridis_d(name="Dispersal\nDistance") + 
    theme(legend.position = c(0.8, 0.8)) + ylab("Exponent") + xlab( expression(~theta)) + facet_wrap(~sim_type)
ggsave("figure/FireNL_exponent_dispersal_theta_season.png",width=8,height=6,units="in",dpi=600)


```

# Compare with data with simulations 

```{r readCompareSimEstimated_bF, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}

#
# Read simulations based on fitted bF 
#

pexp <- readRDS(file.path("Simulations", "fittedExponentsFireNL40years.rds")) %>% mutate(tot_patch= tot_patch/(450*450)*100,max_patch=max_patch/(450*450)*100,num_patch=num_patch/(450*450)*100, year= year(Date), theta = 1/(Fire_probability * Forest_growth), Percent_forest=Percent_forest*100)

#
# Read simulations based on actual data
#

pexp %>% group_by(Fire_probability,forest_dispersal_distance,theta,Date) %>% filter( type == "pl") %>% mutate(forest_dispersal_distance=factor(round(forest_dispersal_distance)),theta = factor(theta)) %>%  ggplot(aes(Date,max_patch,color=theta)) + geom_jitter() +     scale_color_viridis_d() +  ylab("Max Fire Size %") + xlab("")   + facet_wrap( ~forest_dispersal_distance) 

pexp %>% group_by(forest_dispersal_distance,theta,Date) %>% filter( type == "pl") %>% 
  mutate(forest_dispersal_distance=factor(round(forest_dispersal_distance)),theta = factor(theta)) %>%
  ggplot(aes(Date,median_fire_interval,color=theta)) + geom_jitter() + scale_color_viridis_d() +  ylab("Fire return time") + xlab("") + facet_wrap( ~forest_dispersal_distance) 


#
# Forest % 
#
facet_labels <- c("A.", "B.", "C.", "D.")
names(facet_labels) <- c("fixed1", "fixed66", "seasonal1", "seasonal66")
pexp %>% group_by(forest_dispersal_distance,theta,Date) %>% filter( type == "pl") %>%  
      mutate(forest_dispersal_distance=factor(round(forest_dispersal_distance)),theta = factor(theta),cluster= paste0(sim_type,forest_dispersal_distance)) %>%
  ggplot(aes(Date,Percent_forest,color=theta)) + geom_jitter(alpha=0.5) + scale_color_viridis_d(name=expression(~theta)) +  ylab("Forest %") + xlab("") + #+ facet_grid( forest_dispersal_distance ~ sim_type) +
  theme(legend.position = c(0.06, 0.88)) +
  facet_wrap(.~cluster, ncol = 2, labeller = labeller(cluster = as_labeller(facet_labels))) +   
  theme(strip.background = element_rect(fill = "white", color = "white"), strip.text.x = element_text(hjust = 0, size = 12)) 

ggsave("figure/FireNL_ForestPercent_dispersal_theta_season.png",width=8,height=6,units="in",dpi=600)

pexp %>% group_by(forest_dispersal_distance,theta,Date) %>% filter( type == "pl") %>% 
    mutate(forest_dispersal_distance=factor(forest_dispersal_distance),theta = factor(theta)) %>% 
  ggplot(aes(Date,num_patch,color=theta)) + geom_jitter(alpha=0.5) + scale_color_viridis_d() +  ylab("Number of Fires") + xlab("")  + facet_wrap( ~forest_dispersal_distance)  

pexp %>% group_by(forest_dispersal_distance,theta,Date) %>% filter( type == "pl") %>% 
    mutate(forest_dispersal_distance=factor(forest_dispersal_distance),theta = factor(theta)) %>% 
  ggplot(aes(Date,tot_patch,color=theta)) + geom_jitter(alpha=0.5) + scale_color_viridis_d() +  ylab("Total of Fire size") + xlab("")  + facet_wrap( ~forest_dispersal_distance)  


pexp %>% group_by(forest_dispersal_distance,theta,Date) %>% filter( type == "pl") %>% 
  mutate(forest_dispersal_distance=factor(forest_dispersal_distance),theta = factor(theta)) %>% 
  ggplot(aes(theta,num_patch,color=theta)) + geom_violin() + scale_fill_viridis_d(guide=FALSE) +    scale_color_viridis_d(guide=FALSE) +  ylab("Number of Fires") + xlab("")  + facet_wrap( ~forest_dispersal_distance)  

pexp %>% group_by(forest_dispersal_distance,theta) %>% filter( type == "pl", Date==max(Date)) %>%
    mutate(forest_dispersal_distance=factor(forest_dispersal_distance),theta = factor(theta),median_fire_interval=median_fire_interval/365) %>%  ggplot(aes(theta,median_fire_interval,color=theta)) + geom_jitter() + scale_fill_viridis_d(guide=FALSE) +    scale_color_viridis_d(guide=FALSE) +  ylab("Fire Return Time") + xlab("")  + facet_wrap( ~forest_dispersal_distance)  



# Number of fires vs Total size
#
pexp %>% group_by(forest_dispersal_distance,theta,Date) %>% filter( type == "pl", theta==2500) %>%  
      mutate(forest_dispersal_distance=factor(round(forest_dispersal_distance)),theta = factor(theta)) %>% 
      ggplot(  aes(num_patch,tot_patch, color=theta ))+ geom_point() + theme_bw() +
      scale_color_viridis_d()  + 
      xlab("Number of fires (%)") +  
      ylab("Total Size (%)") +  
      theme(axis.text.x = element_text(angle = 50, hjust = 1))  + facet_grid( sim_type~forest_dispersal_distance)  + theme(legend.justification=c(1,0), legend.position=c(.99,0.01))


# Max fire vs Total size 
#
#

facet_labels <- c("A.", "B.", "C.", "D.")
names(facet_labels) <- c("fixed1", "fixed66", "seasonal1", "seasonal66")
pexp %>% group_by(forest_dispersal_distance,theta,Date) %>% filter( type == "pl") %>%  
      mutate(forest_dispersal_distance=factor(round(forest_dispersal_distance)),theta = factor(theta),cluster= paste0(sim_type,forest_dispersal_distance)) %>%  
      ggplot(  aes(max_patch,tot_patch, color=theta ))+ geom_point(alpha=0.5) + theme_bw() +
      scale_color_viridis_d(name=expression(~theta)) + 
      xlab("Maximum fire size (%)") +  
      ylab("Total Size (%)") +  
      theme(axis.text.x = element_text(angle = 0, hjust = 1))  +  
      theme(legend.justification=c(1,0), legend.position=c(.99,0.01)) +
      facet_wrap(.~cluster, ncol = 2, labeller = labeller(cluster = as_labeller(facet_labels))) +   
        theme(strip.background = element_rect(fill = "white", color = "white"), strip.text.x = element_text(hjust = 0, size = 12)) 
ggsave("figure/FireNL_TotSizeVsMaxSize_dispersal_theta_season.png",width=8,height=6,units="in",dpi=600)



# Forest % vs Total size
#
pexp %>% group_by(forest_dispersal_distance,theta,Date) %>% filter( type == "pl") %>%  
      mutate(forest_dispersal_distance=factor(forest_dispersal_distance),theta = factor(theta)) %>% 
      ggplot(  aes(Percent_forest,tot_patch, color=theta ))+ geom_point(alpha=0.5) + theme_bw() +
      scale_color_viridis_d()  + 
      xlab("Forest (%)") +  
      ylab("Total Size (%)") +  
      theme(axis.text.x = element_text(angle = 90, hjust = 1))  + facet_wrap( ~forest_dispersal_distance)  + theme(legend.justification=c(1,0), legend.position=c(.99,0.01))


# Forest % vs Max size
#
pexp %>% group_by(forest_dispersal_distance,theta,Date) %>% filter( type == "pl") %>%  
      mutate(forest_dispersal_distance=factor(forest_dispersal_distance),theta = factor(theta)) %>% 
      ggplot(  aes(Percent_forest,max_patch, color=theta ))+ geom_point(alpha=0.5) + theme_bw() +
      scale_color_viridis_d()  + 
      xlab("Forest (%)") +  
      ylab("Max Size (%)") +  
      theme(axis.text.x = element_text(angle = 90, hjust = 1))  + facet_grid( sim_type~forest_dispersal_distance)   + theme(legend.justification=c(1,0), legend.position=c(.99,0.01))

#
# Mean
#

pexp %>% group_by(forest_dispersal_distance,theta,Date) %>% filter( type == "pl") %>% mutate(forest_dispersal_distance=factor(forest_dispersal_distance),theta=factor(theta)) %>% summarise(max_patch=mean(max_patch)) %>% ggplot(aes(Date,max_patch,color=theta)) + geom_point()  +  scale_color_viridis_d() +  ylab("Max Fire Size %") + xlab("")   + facet_wrap( ~forest_dispersal_distance) 

pexp %>% group_by(forest_dispersal_distance,theta,Date) %>% filter( type == "pl") %>% mutate(forest_dispersal_distance=factor(forest_dispersal_distance),theta=factor(theta)) %>% summarise(tot_patch=mean(tot_patch)) %>% ggplot(aes(Date,tot_patch,color=theta)) + geom_point()  +  scale_color_viridis_d() +  ylab("Total Fire Size") + xlab("") + facet_wrap( ~forest_dispersal_distance) 
  

pexp %>% group_by(forest_dispersal_distance,theta,Date) %>% filter( type == "pl") %>%
  mutate(forest_dispersal_distance=factor(forest_dispersal_distance),theta=factor(theta))  %>% 
  summarise(Percent_forest=mean(Percent_forest)) %>% ggplot(aes(Date,Percent_forest,color=theta)) + geom_point() + scale_fill_viridis_d(guide=FALSE) +    scale_color_viridis_d() +  ylab("Forest %") + xlab("")  + facet_wrap( ~forest_dispersal_distance) 

pexp %>% group_by(forest_dispersal_distance,theta,Date) %>% filter( type == "pl") %>% mutate(forest_dispersal_distance=factor(forest_dispersal_distance)) %>% summarise(num_patch=mean(num_patch)) %>% ggplot(aes(Date,num_patch,color=forest_dispersal_distance)) + geom_point() + scale_fill_viridis_d(guide=FALSE) +    scale_color_viridis_d(guide=FALSE) +  ylab("Number of Fires") + xlab("")  



#
# Save plot with simulations spatial distribution  
#
require(png)
require(cowplot)
p1 <- grid::rasterGrob(readPNG("figure/FireNL_450_theta2500_noseason_de102.png"))
p2 <- grid::rasterGrob(readPNG("figure/FireNL_450_theta2500_noseason_de20155.png"))
p3 <- grid::rasterGrob(readPNG("figure/FireNL_450_theta25_noseason_de102.png"))
p4 <- grid::rasterGrob(readPNG("figure/FireNL_450_theta25_noseason_de20155.png"))

prow <- plot_grid(
  p1,p2,p3,p4,
  align = 'vh',
  nrow= 2, labels=c("A.", "B.", "C.", "D.",label_size = 12), scale=c(.95,.95,.95,.95)
  )

save_plot("figure/FireNL_450_theta25-2500_de102-20155_noseason.png",prow,base_width=8,base_height=6,dpi=600)


p1 <- grid::rasterGrob(readPNG("figure/FireNL_450_theta2500_season90_de102.png"))
p2 <- grid::rasterGrob(readPNG("figure/FireNL_450_theta2500_season90_de20155.png"))
p3 <- grid::rasterGrob(readPNG("figure/FireNL_450_theta25_season90_de102.png"))
p4 <- grid::rasterGrob(readPNG("figure/FireNL_450_theta25_season90_de20155.png"))

prow <- plot_grid(
  p1,p2,p3,p4,
  align = 'vh',
  nrow= 2, labels=c("A.", "B.", "C.", "D.",label_size = 12), scale=c(.95,.95,.95,.95)
  )

save_plot("figure/FireNL_450_theta25-2500_de102-20155_season90.png",prow,base_width=8,base_height=6,dpi=600)


prow
```

